\section{Kan's loop group construction}

Recall that simplicial sets form a combinatorial model for the homotopy theory of spaces.
This may be expressed using the language of model categories by stating that the adjoint functors
\[
\bars{\ \cdot\ } \colon \sSet \leftrightarrows \mathsf{Top} :\! \sSing
\]
defines a Quillen equivalence between model category structures, where $\Top$ is equipped with the classical \textit{Quillen model structure} and $\sSet$ with the \textit{Kan--Quillen model structure}.
We refer to the induced model structure on $\sSet^0$ by the same name.

The based loop space of a reduced simplicial set may be modeled combinatorially through the \textit{Kan loop group functor}
\[
\kan \colon \sSet^0 \to \sGrp
\]
defined as follows.
For any reduced simplicial set $X$ each $\kan(X)_n$ is the group with one generator $\overline{x}$ for every simplex $x \in X_{n+1}$ modulo the relation $\overline{s_0(x)} = e$, where $e$ denotes the identity element of the group.

The face and degeneracy maps are defined by the following equations:
\[
\delta_0(\overline{x}) = \overline{\delta_1(x)} \cdot (\overline{\delta_0(x)})^{-1}, \qquad \quad
\delta_n(\overline{x}) = \overline{\delta_{n+1}(x)} \text{ for } n>0, \qquad \quad
s_n(\overline{x}) = \overline{s_{n+1}(x)}.
\]

Berger showed that this construction models the based loop space as a topological monoid.
More precisely, he showed that for any reduced simplicial set $X$, the geometric realization of the Kan loop group $\bars{\kan(X)}$ is weak homotopy equivalent, as a topological monoid, to the based loop space $\loops \bars{X}$ on the geometric realization of $X$ \cite{berger1995loops}.

Using Kan's ideas, Quillen constructed a model category structure on the category $\sGrp$ of simplicial groups determined by
weak equivalences and fibrations being maps of simplicial groups whose underlying maps of simplicial sets are weak homotopy equivalences and Kan fibrations, respectively.
We call this model category structure the \textit{Kan model structure} on $\sGrp$.

The Kan loop group functor has a right adjoint $\classifying \colon \sGrp \to \sSet^0$ modeling the classifying space of a simplicial group.
As proven for example in \cite[Chapter V]{goerss2009simplicial} we have the following.

\begin{proposition} \label{p:kan adjuntion}
	The adjunction
	\[
	\kan \colon \sSet^0 \leftrightarrows \sGrp :\! \classifying
	\]
	defines a Quillen equivalence when $\sSet^0$ and $\sGrp$ are equipped respectively with the Kan--Quillen and the Kan model category structures.
\end{proposition}