% !TEX root = ../cobar.tex

\section{Conventions and preliminaries} \label{s:preliminaries}

\subsection{Presheaves}

Given categories $\sB$ and $\sC$ we denote their associated \textit{functor category} by $\Fun[\sB, \sC]$.
Recall that a category is said to be \textit{small} if its objects and morphisms form sets.
We denote the category of small categories by $\Cat$.
A category is said to be \textit{cocomplete} if any functor to it from a small category has a colimit.
If $\sA$ is small and $\sC$ cocomplete, then the (left) \textit{Kan extension of $g$ along $f$} exists for any pair of functors $f$ and $g$ in the diagram below, and it is the initial object in $\Fun[\sB, \sC]$ making
\begin{equation*}
	\begin{tikzcd}[column sep=normal, row sep=normal]
		\sA \arrow[d, "f"'] \arrow[r, "g"] & \sC \\
		\sB \arrow[dashed, ur, bend right] & \quad
	\end{tikzcd}
\end{equation*}
commute.
A Kan extension along the \textit{Yoneda embedding}, i.e., the functor
\[
\yoneda \colon \sA \to \Fun[\sA^\op, \Set]
\]
induced by the assignment
\[
a \mapsto \big( a^\prime \mapsto \sA(a^\prime, a) \big),
\]
is referred to as a \textit{Yoneda extension}.
We refer to the objects in $\Fun[\sA^\op, \Set]$ as \textit{presheaves on $\sA$} and to those in the image of the Yoneda embedding as \textit{representables}.
We remark that for any presheaf $P$ on $\sA$ and object $a$ of $\sA$
\[
P \, \cong \colim_{\yoneda(a) \to P} \yoneda(a).
\]
In other words, any presheaf is isomorphic to a colimit of representables.

\subsection{Chain complexes}

Throughout this article $\k$ denotes a commutative and unital ring and we work over its associated closed symmetric monoidal category of differential (homologically) graded $\k$-modules $(\Ch, \otimes, \k)$.
We refer to the objects and morphisms of this category as \textit{chain complexes} and \textit{chain maps} respectively.
We denote by $\Hom(C, C^\prime)$ the chain complex of $\k$-linear maps between chain complexes $C$ and $C^\prime$, and refer to the functor $\Hom(-, \k)$ as \textit{linear duality}.
The $i^\th$ \textit{suspension} functor $s^i \colon \Ch \to \Ch$ is defined at the level of graded modules by $(s^{i}M)_n = M_{n-i}$.

\subsection{Coalgebras} \label{ss:coalgebras}

A \textit{coalgebra} consists of a chain complex $C$ and chain maps $\Delta \colon C \to C \otimes C$ and $\varepsilon \colon C \to \k$ satisfying the usual coassociativity and counitality relations.
This notion is equivalent to that of a comonoid in $\Ch$.
Denote by $\coAlg$ the category of coalgebras with morphisms being structure preserving chain maps.
The category $\coAlg$ is symmetric monoidal, with braiding induced from $\Ch$ and structure maps of a product $C \otimes C^\prime$ given by
\begin{gather*}
	C \otimes C^\prime \xra{\Delta \otimes \Delta^\prime}
	(C \otimes C) \otimes (C^\prime \otimes C^\prime) \xra{(23)}
	(C \otimes C^\prime) \otimes (C \otimes C^\prime), \\
	C \otimes C^\prime \xra{\varepsilon \otimes \varepsilon^\prime}
	\k \otimes \k \xra{\cong} \k.
\end{gather*}

\subsection{Monoids}

A \textit{monoid} $(M, \mu, \eta)$ in a monoidal category $(\sC, \otimes, \mathbb{1})$ is an object $M$ together with two morphisms $\mu \colon M \otimes M \to M$, called \textit{multiplication} and $\eta \colon \mathbb{1} \to M$, called \textit{unit}, satisfying the usual associativity and unital relations.
Monoids in $\Ch$ are called \textit{algebras} and those in $\coAlg$ are called \textit{bialgebras}.
We denote the category of monoids in $\sC$ by $\Mon_{\sC}$ and remark that a monoidal functor $\sC \to \sC^\prime$ induces a functor between their categories of monoids $\Mon_\sC \to \Mon_{\sC^\prime}$.

\subsection{The cobar construction} \label{ss:cobar construction}

A \textit{coaugmentation} on a coalgebra $C$ is a morphism $\nu \colon \k \to C$ in $\coAlg$.
A coalgebra is said to be \textit{coaugmented} if it is equipped with a coaugmentation.

We denote by $\coAlg^\ast$ the category of coaugmented coalgebras with morphisms being coaugmentation preserving coalgebra morphisms.

The \textit{cobar construction} is the functor
\[
\cobar \colon \coAlg^\ast \to \Mon_{\Ch}
\]
defined on objects as follows.
Let $(C, \partial, \Delta, \varepsilon, \nu)$ be the data of a coaugmented coalgebra.
Denote by $\overline{C}$ the cokernel of $\nu \colon \k \to C$ and recall that $s^{-1}$ is the $(-1)^\th$ suspension.
The cobar construction $\cobar C$ of this coaugmented coalgebra is the graded module
\[
T(s^{-1} \overline{C}) = \k \oplus s^{-1}\overline{C} \oplus (s^{-1}\overline{C})^{\otimes 2} \oplus (s^{-1}\overline{C})^{\otimes 3} \oplus \dots
\]
regarded as a monoid in $\Ch$ with $\mu \colon T(s^{-1} \overline{C})^{\otimes 2} \to T(s^{-1} \overline{C})$ given by concatenation, $\eta \colon \k \to T(s^{-1} \overline{C})$ by the obvious inclusion, and differential constructed by extending the linear map
\[
- s^{-1} \circ \partial \circ s^{+1} \ + \ (s^{-1} \otimes s^{-1}) \circ \Delta \circ s^{+1} \colon
s^{-1} \overline{C} \to s^{-1}\overline{C} \oplus (s^{-1}\overline{C} \otimes s^{-1}\overline{C}) \hookrightarrow T(s^{-1}C)
\]
as a derivation.

In this article we are mostly concerned with \textit{connected} coalgebras, namely coaugmented coalgebras that are non-negatively graded and for which the coaugmentation induces an isomorphism $\k \cong C_0$ of $\k$-modules.
We denote by $\coAlg^0$ the full subcategory of $\coAlg^\ast$ consisting of connected coalgebras, and remark that if $C \in \coAlg^0$, then $ \cobar(C)$ is concentrated on non-negative degrees.

The object $ \cobar(C) \in \Mon_{\Ch}$ may be equipped with a natural augmentation map $\cobar(C) \to \k$ in $\Mon_{\Ch}$ given by the projection map to the first summand term.
Therefore, the cobar construction may be regarded as a functor from coaugmented coalgebras to augmented algebras.
However, we will not require the data of augmentations in this article.

We now review the central examples of coaugmented coalgebras to which Adams applied the cobar construction.

\subsection{Simplicial sets}

The \textit{simplex category} is denoted by $\simplex$ and its objects by $[n]$.
Its morphisms are generated by the usual (simplicial) \textit{coface} and \textit{codegeneracy maps}
\[
\delta_i \colon [n-1] \to [n], \qquad \sigma_i \colon [n+1] \to [n]
\]
for $j \in \{0, \dots, n\}$.
The category of \textit{simplicial sets} $\Fun[\simplex^\op, \Set]$ is denoted by $\sSet$ and the \textit{standard $n$-simplex} $\yoneda [n] = \simplex(-, [n])$ by $\simplex^n$.
For any simplicial set $X$ we write, as usual, $X_n$ instead of $X \big( [n] \big)$, and identify the elements of $\simplex^n_m$ with increasing tuples $[v_0, \dots, v_m]$ where $v_i \in \{0, \dots, n\}$.

If $X$ is such that $X_0$ is a singleton set we say that $X$ is \textit{reduced}.
We denote the full subcategory of reduced simplicial sets by $\sSet^0$.

\subsection{Simplicial chains} \label{ss:simplicial sets}

For non-negative integers $m$ and $n$, let $\simplex_{\deg} \big( [m], [n] \big)$ be the subset of \textit{degenerate morphisms} in $\simplex \big( [m], [n] \big)$, i.e., those of the form $\sigma_i \circ \tau$ with $\tau$ any morphism in $\simplex \big( [m], [n+1] \big)$.
The functor of \textit{simplicial chains} $\schains \colon \sSet \to \Ch$ is the Yoneda extension of the functor $\simplex \to \Ch$ defined next.
It assigns to an object $[n]$ the chain complex having in degree $m$ the module
\[
\frac{\k\{\simplex \big( [m], [n] \big) \}}{\k\{\simplex_{\deg} \big( [m], [n] \big) \}},
\]
and differential
\[
\partial(\tau) = \sum (-1)^i \tau \circ \delta_i.
\]
To a morphism $\tau \colon [n] \to [n^\prime]$ it assigns the chain map
\[
\begin{tikzcd}[row sep=-3pt, column sep=normal,
	/tikz/column 1/.append style={anchor=base east},
	/tikz/column 2/.append style={anchor=base west}]
	\schains(\simplex^n)_m \arrow[r] & \schains(\simplex^{n^\prime})_m \\
	\big( [m] \to [n] \big) \arrow[r, mapsto] & \big( [m] \to [n] \xra{\tau} [n^\prime] \big).
\end{tikzcd}
\]
When no confusion arises from doing so we write $\chains$ instead of $\schains$ and refer to it simply as the functor of \textit{chains}.

\subsection{Alexander--Whitney coalgebra} \label{ss:aw coalgebra}

We review a classical lift to coalgebras of the functor of chains:
\[
\begin{tikzcd}
	& \coAlg \arrow[d] \\
	\sSet \arrow[r, "\schains"] \arrow[ur, "\schainsA", out=70, in=180] & \Ch.
\end{tikzcd}
\]

Using a Yoneda extension, it suffices to equip the chains on standard simplices with a natural coalgebra structure.
For any $n \in \N$, define $\epsilon \colon \chains(\simplex^n) \to \k$ by
\[
\epsilon \big( [v_0, \dots, v_q] \big) = \begin{cases} 1 & \text{ if } q = 0, \\ 0 & \text{ if } q>0, \end{cases}
\]
and $\Delta \colon \chains(\simplex^n) \to \chains(\simplex^n)^{\otimes2}$ by
\[
\Delta \big( [v_0, \dots, v_q] \big) = \sum_{i=0}^q \ [v_0, \dots, v_i] \otimes [v_i, \dots, v_q].
\]

If $X$ is a reduced simplicial set then $\schainsA(X)$ becomes a connected coalgebra with the coaugmentation $\nu \colon \k \to \schains(X)$ induced by sending $1$ to the basis element represented by the unique $0$-simplex of $X$.
Hence, the functor $\schainsA$ restricts to a functor
\[
\schainsA \colon \sSet^0 \to \coAlg^0.
\]

\subsection{Simplicial singular complex}

Consider the topological $n$-simplex
\[
\gsimplex^{n} = \{(x_0, \dots, x_n) \in [0,1]^{n+1} \mid \ \textstyle{\sum}_i x_i = 1\}.
\]
The assignment $[n] \to \gsimplex^n$ defines a functor $\simplex \to \Top$ whose Yoneda extension is known as (simplicial) \textit{geometric realization}.
It has a right adjoint $\sSing \colon \Top \to \sSet$ given by
\[
Z \to \Big( [n] \mapsto \Top(\gsimplex^n, Z) \Big)
\]
and referred to as the \textit{simplicial singular complex} of the topological space $Z$.
The chain complex $\chains(\cSing Z)$ is referred to as the \textit{simplicial singular chains} of $Z$.

We modify this construction for a pointed topological space $(Z,z)$ by only considering maps $\gsimplex^n \to Z$ sending all vertices to $z$.
This produces a reduced simplicial set $\sSing(Z,z)$.
We will refer to $\schains(\sSing(Z,z))$ as the \textit{simplicial singular chains on $(Z,z)$} and for simplicity denote it by $\sSchains(Z,z)$.

\subsection{Cubical sets} \label{ss:cubical sets}

The objects of the \textit{cube category} $\cube$ are the sets $2^n = \{0, 1\}^n$ with $2^0 = \{0\}$ for $n \in \N$, and its morphisms are generated by the \textit{coface, codegeneracy} and \textit{coconnection} maps defined by
\begin{align*}
	\delta_i^\varepsilon & =
	\mathrm{id}_{2^{i-1}} \times \delta^\varepsilon \times \mathrm{id}_{2^{n-1-i}} \colon 2^{n-1} \to 2^n, \\
	\sigma_i & =
	\mathrm{id}_{2^{i-1}} \times \, \sigma \times \mathrm{id}_{2^{n-i}} \quad \colon 2^{n} \to 2^{n-1}, \\
	\gamma_i & =
	\mathrm{id}_{2^{i-1}} \times \, \gamma \times \mathrm{id}_{2^{n-i}} \quad \colon 2^{n+1} \to 2^{n},
\end{align*}
where $\varepsilon \in \{0,1\}$ and
\[
\begin{tikzcd}
	1 \arrow[r, out=45, in=135, "\delta^0"] \arrow[r, out=-45, in=-135, "\delta^1"'] & 2 \arrow[l, "\sigma"'] &[-10pt] \arrow[l, "\gamma"'] 2 \times 2
\end{tikzcd}
\]
are defined by
\begin{gather*}
	\delta^0(0) = 0, \qquad
	\delta^1(0) = 1, \qquad
	\sigma(0) = \sigma(1) = 0, \qquad \\
	\gamma^1(0,0) = 0, \qquad
	\gamma^1(0,1) = 1, \qquad
	\gamma^1(1,0) = 1, \qquad
	\gamma^1(1,1) = 1.
\end{gather*}
We remark that the cube category without connections will not be considered in this work.

The category of \textit{cubical sets} $\Fun[\cube^\op, \Set]$ is denoted by $\cSet$ and
the \textit{standard $n$-cube} $\yoneda(2^n) = \cube(-, 2^n)$ by $\cube^n$.
For any cubical set $Y$ we write, as usual, $Y_n$ instead of $Y(2^n)$.

\subsection{Cubical chains}

For non-negative integers $m$ and $n$, let $\cube_{\deg}(2^m, 2^n)$ be the subset of \textit{degenerate morphisms} in $\cube(2^m, 2^n)$, i.e., those of the form $\sigma_i \circ \tau$ or $\gamma_i \circ \tau$ with $\tau$ any morphism in $\cube(2^m, 2^{n+1})$.
The functor of (cubical) \textit{chains} $\cchains \colon \cSet \to \Ch$ is the Yoneda extension of the functor $\cube \to \Ch$ defined next.
It assigns to an object $2^n$ the chain complex having in degree $m$ the module
\[
\frac{\k\{\cube(2^m, 2^n)\}}{\k\{\cube_{\deg}(2^m, 2^n)\}}
\]
and differential induced by
\[
\partial (\id_{2^n}) = \sum_{i=1}^{n} \ (-1)^i \
\big(\delta_i^1 - \delta_i^0 \big).
\]
To a morphism $\tau \colon 2^n \to 2^{n^\prime}$ it assigns the chain map
\[
\begin{tikzcd}[row sep=-3pt, column sep=normal,
	/tikz/column 1/.append style={anchor=base east},
	/tikz/column 2/.append style={anchor=base west}]
	\cchains(\cube^n)_m \arrow[r] & \cchains(\cube^{n^\prime})_m \\
	\big( 2^m \to 2^n \big) \arrow[r, mapsto] & \big( 2^m \to 2^n \stackrel{\tau}{\to} 2^{n^\prime} \big).
\end{tikzcd}
\]

When no confusion arises we write $\chains$ instead of $\cchains$.

We remark that $\chains(\cube^n)$ is canonically isomorphic to $\chains(\cube^1)^{\otimes n}$ and that $\chains(\cube^1)$ is canonically isomorphic to the cellular chains on the topological interval with its usual CW structure.
We use these isomorphisms to denote the elements of $\chains(\cube^n)$ as sums of terms of the form $x_1 \otimes \dots \otimes x_n$ with $x_i \in \big\{[0], [0,1], [1] \big\}$.

\subsection{Serre coalgebra} \label{ss:serre coalgebra}

We review a classical lift of the functor of cubical chains to the category of coalgebras:
\[
\begin{tikzcd}
	& \coAlg \arrow[d] \\
	\cSet \arrow[r, "\cchains"] \arrow[ur, "\cchainsA", out=70, in=180] & \Ch.
\end{tikzcd}
\]
Using a Yoneda extension, it suffices to equip the chains on standard cubes with a natural coalgebra structure.
For any $n \in \N$, define $\epsilon \colon \chains(\cube^n) \to \k$ by
\[
\epsilon \left( x_1 \otimes \dots \otimes x_d \right) = \epsilon(x_1) \dotsm \epsilon(x_n),
\]
where
\[
\epsilon([0]) = \epsilon([1]) = 1, \qquad \epsilon([0, 1]) = 0,
\]
and $\Delta \colon \chains(\cube^n) \to \chains(\cube^n)^{\otimes 2}$ by
\[
\Delta (x_1 \otimes \dots \otimes x_n) =
\sum \pm \left( x_1^{(1)} \otimes \dots \otimes x_n^{(1)} \right) \otimes
\left( x_1^{(2)} \otimes \dots \otimes x_n^{(2)} \right),
\]
where the sign is determined using the Koszul convention, and we are using Sweedler's notation
\[
\Delta(x_i) = \sum x_i^{(1)} \otimes x_i^{(2)}
\]
for the chain map $\Delta \colon \chains(\cube^1) \to \chains(\cube^1)^{\otimes 2}$ defined by
\[
\Delta([0]) = [0] \otimes [0], \quad \Delta([1]) = [1] \otimes [1], \quad \Delta([0,1]) = [0] \otimes [0,1] + [0,1] \otimes [1].
\]

Using the canonical isomorphism $\chains(\cube^n) \cong \chains(\cube^1)^{\otimes n}$, the coproduct $\Delta$ can be described as the composition
\[
\begin{tikzcd}
	\chains(\cube^1)^{\otimes n} \arrow[r, "\Delta^{\otimes n}"] & \left( \chains(\cube^1)^{\otimes 2} \right)^{\otimes n} \arrow[r, "sh"] & \left( \chains(\cube^1)^{\otimes n} \right)^{\otimes 2},
\end{tikzcd}
\]
where $sh$ is the shuffle map that places tensor factors in odd position first.

\subsection{Cubical singular complex}

Consider the topological $n$-cube
\[
\gcube^{n} = \{(x_1, \dots, x_n) \mid x_i \in [0,1]\}.
\]
The assignment $2^n \to \gcube^n$ defines a functor $\cube \to \Top$ whose Yoneda extension is known as \textit{geometric realization}.
It has a right adjoint $\cSing \colon \Top \to \cSet$ given by
\[
Z \to \Big(2^n \mapsto \Top(\gcube^n, Z)\Big)
\]
and referred to as the \textit{cubical singular complex} of the topological space $Z$.
We will refer to $\chains(\cSing Z)$ as the \textit{cubical singular chains} of $Z$ and for simplicity denote it by $\cSchains(Z)$.